---
format:
    html:
        df-print: tibble
---

# Introduction {#sec-introduction .unnumbered}

The [Civil Service People Survey][cspshub] is is the annual employee attitudes
survey that has carried out across the UK Civil Service since 2009. It is an
important management tool for managers and leaders working within in the Civil
Service. It is also a useful tool for external parties that hold senior
officials and government ministers to account for the leadership and management
of the Civil Service.

This introduction covers: [who this companion is for](#sec-intro-who-for); a
brief explanation of
[how the People Survey is run](#sec-intro-how-survey-is-run); key
[underlying concepts](#sec-survey-concepts) in the structure of survey's data;
and, an outline of the harmonised
[concepts and identifiers](#sec-intro-harmonised-concepts) and the
[structure](#sec-intro-harmonised-data) of the harmonised datasets.

This companion uses the terms 'Civil Service People Survey' and 'People Survey'
interchangeably; the acronym 'CSPS' is used sparingly, usually were space is
at a premium. A more complete guide to [terminology](01_terminology.qmd) is
also available.

## Who this companion is for {#sec-intro-who-for}

This companion mainly acts as documentation for the data processing in the
`csps-data` repository. It has been called a 'companion' because it takes a
narrative approach to documentation that also includes wider discussion of the
survey's background and operation, and explains the rationale for behind
decisions taken to harmonise the data over time.

The main audience for this companion is those who wish to reuse the harmonised
data for their own analysis. It also serves as a record of decisions made in
regards to harmonising aspects of the survey's data structure that change over
time. This companion may also be useful for those working inside the Civil
Service that are working with multi-year data and/or interested in comparing
results between organisations or to Civil Service benchmarks.

## How the People Survey is run {#sec-intro-how-survey-is-run}

The Civil Service People Survey is coordinated by the
[People Survey Team][cspshub] based in the Cabinet Office, they are responsible
for the overall management and operation of the survey and publishing of the
results. The survey's fieldwork and internal reporting is provided by an
external third-party contractor, currently
[Qualtrics](https://www.qualtrics.com/). Alongside the central team and
external contractor, participating organisations have a 'survey manager' that
acts as the liaison point for each organisation, survey managers are responsible
for collating crucial 'local' information necessary for the running of the
survey (such as the internal team hierarchy), publicising the survey and
encouraging participation during the fieldwork period, and disseminating the
results within their organisation.

## The need for a harmonised dataset {#sec-intro-need}

While data and results from the survey have been published each year, there is
no single dataset or API that makes it easy to access the data. The range of
data published each year varies, as do aspects of the survey content and
coverage.

Although there is some degree of consistency and stability in the survey's
questionnaire over time there are additions and removals of questions that
have occurred over the past 16 years. Organisational coverage varies over time
as organisations are established, merged, abolished or changed in other ways.
The range of demographic categories either asked about or published has also
changed over time.

These challenges prevent easy re-use of the data, including by individuals
within the UK Civil Service. The aim of the data processing documented in this
companion is to provide a public resource that makes it easy for others to
re-use the data.

## Original survey data concepts {#sec-survey-concepts}

Before any processing we must first define core concepts in the survey's data
structures and methodology.

### Attitudinal questions and measures

The main body of the survey's questionnaire is made up of individual questions
that measure individual's attitudes and experiences of working in their civil
service organisation over the past year. These questions are often rated through
a five-point Likert scale of 'strongly disagree' to 'strongly agree', see the
section on [response categories](#sec-intro-response-categories) for more
details. Most of these attitudinal questions are also used to derive composite
measures that are used as headline results for the survey, see the section on
[indexes and theme scores](#sec-intro-composite-measures) for more details.
In the context of the data processing, together, these are referred to as
attitudinal *questions and measures*. In programming code the term `qm` has
been typically used to refer to these data. See !!! for more details.

### Organisations

Respondents to the survey participate as part of an *organisation*, typically
(but not always) the government department or agency they work for. Questions
in the survey typically ask respondents to consider their responses in relation
to this organisational unit. In programming code the term `org` has typically
been used to refer to organisations, and `dept_group` is used to refer to groups
of organisations related to a specific government department. See !!! for more
details.

### Civil Service benchmark and all respondents results

The headline cross-government results from the survey are the
*Civil Service benchmark* results, this is calculated for each question and
measure as the median score of participating organisations.

The results for all respondents to the survey are also published but these
scores are largely determined by the experiences of respondents based in the
5-6 largest civil service organisations. These scores may be referred to as
either *all respondents*, *all civil servants* or *mean scores*.

### Demographic questions

Alongside the attitudinal questions the People Survey asks a range of
*demographic* questions relating to both their job and personal
characteristics that are used to understand the experiences of different groups
of civil servants. The demographics comprise both questions and categories, in
programming code the term `demog` is typically used to refer to these two
aspects collectively, `demq` for demographic questions, `cat` for categories,
and `demcat` for combinations of questions and categories (since categories are
not always unique across the questions).

In addition to the results for all respondents split by demographic questions
and categories there are also more detailed results published for five sets of
demographics (sex/gender, ethnicity, health status, sexual orientation and
socio-economic background). These detailed demographic results publications
provide results of cross-tabulation of these five demographics with all other
demographic questions, for example results for men and women by grade. The
detailed demographic results also include a summary of organisation scores for
each of the five demographics.

### Typology of data types {#sec-intro-typology}

From the structure of the People Survey data described above we can devise a
simple taxonomy of the different 'types' of data published by the People Survey.
The attitudinal questions and measures are excluded from the taxonomy, with the
taxonomy acting as a guide to the type of units of analysis covered by different
sets of attitudinal question and measure results.

* The **benchmark results** - the 'Civil Service benchmark', i.e. the median
  scores of participating organisations.
* The **mean scores** - the results for all respondents to the People Survey.
* The **organisation results** - the overall results for each organisation.
* The **demographic results** - the results for all respondents by individual
  demographic questions and categories.
* The **detailed demographic results** - results for multiple combinations of
  demographic questions and categories.
* The **organisation demographic results** - results of the People Survey's
  headline measures for each organisation and select demographic categories.

### Response categories {#sec-intro-response-categories}

The majority of attitudinal questions are asked on a five-point scale ranging
from 'strongly disagree' to 'strongly agree'. For these questions the
survey results present the 'percent positive' which reflects the proportion
answering 'agree' or 'strongly agree' [^pdf_reports].

In addition to the five-point agreement scale the following scales are used for
some questions:

* *Yes/no scales*: offering respondents either a simple binary 'yes' 'no'
  choice or sometimes with a 'don't know' or 'prefer not to say' option.
* *Temporal scales*: offering respondents options to state how often a
  situation occurs, typically ranging from 'always' to 'never' but also
  including scales that refer to specific time periods (e.g. 'weekly' or
  'monthly').
* *Quality scale*: used for ratings of mental and physical health, ranging
  from 'excellent' to 'very poor'.
* *Numeric extent scale*: used for the personal wellbeing questions,
  respondents are asked to provide a rating from '0' to '10' where a response
  of '0' represents 'not at all' and '10' represents 'completely' (e.g. 'not at
  all satisfied' or 'completely satisfied').
* *Productivity range scale*: used for a question on self-assessed
  productivity, respondents asked how productive they feel they have been
  recently and are given a set of percentage ranges to select from.
* *Stay/leave scale*: used for the question on future intentions, respondents
  are asked about the future intentions of working for their organisation with
  responses ranging from leaving as soon as possible to staying for at least
  the next three years.
* *Multiple choice scales*: offering respondents the ability to select
  multiple categories in response to a question (e.g. to describe the type of
  bullying or harassment the individual has experienced).

### Indexes and theme scores {#sec-intro-composite-measures}

The responses to individual questions are used to calculate a small number of
summary measures, which are used as headline results for the survey. These
summary measures are either called *indexes* when they make use of the full
range of the input questions' response scales or *theme scores* when they
make use of only part of the range of the input questions' response scales. 

* The employee *engagement index* is derived from five questions relating to
  respondents levels' of advocacy, attachment and motivation.
* The *PERMA index* and *Proxy Stress Index* are wellbeing indexes that
  assess the extent to which respondents are 'flourishing' (for the PERMA Index)
  or experiencing factors that potentially increase stress levels (for the
  Proxy Stress Index).
* The nine *theme scores* are derived from the main section of attitudinal
  questions and measure factors that influence employee engagement. These scores
  are calculated relating to the of 'strongly agree' and 'agree' responses to
  each question within each theme.

[^pdf_reports]: Prior to 2020 some organisations published PDF reports of
their results that also included a breakdown of the Likert scale for each
question. A change in contractor for the People Survey ended the production of
these PDF reports and since then most organisations have stopped publishing
their own People Survey results independently of the combined results published
by the Cabinet Office.

## Harmonised survey data concepts {#sec-intro-harmonised-concepts}

The processing documented in this companion aligns the original survey data
published by the Cabinet Office into a set of harmonised datasets that allow
for easy cross-sectional and time-series analysis of the data.

### Unique identifiers

At the core of the processing is the use of
[regular expressions](https://en.wikipedia.org/wiki/Regular_expression) (or
*'regexes'*) which are used to match question or measure text, organisation
names, demographic questions and categories to a set of *unique identifiers*.
There are three sets of unique identifiers:

* **Question and measure identifiers:**
  * `uid_qm_num`: A numeric identifier in the format `0.00.000.00`
  * `uid_qm_txt`: A human-readable text identifier in the format `thm.question`
    (where `thm` is a short-code for the survey section and `question` is short
    phrase to describe the question or measure)
* **Organisation identifiers:**
  * `uid_org_txt`: Typically a 6-letter code to refer to organisations,
    in some cases a shorter 3- to 5-letter code is used
* **Demographic identifiers:**
  * `uid_demq_txt`: Typically a 6-letter code to refer to demographic
    questions, in two cases a 5-letter code is used
  * `uid_cat_txt`: A 6- to 12-character alphanumeric code to refer to
    demographic categories
  * `uid_demcat_num`: A numeric identifier in the format `0.00.00.000` used to
    refer to unique combinations of demographic questions and categories
  * `uid_demcat_txt`: A (somewhat) human-readable text identifier used to
    refer to unique combinations of demographic questions and categories in
    the format `DEMQ_DCAT` where `DEMQ` refers to the demographic question
    and `DCAT` to the category.

### Output data files {#sec-intro-harmonised-data}

The output data files are a combination of 'datasets' and 'lookups',
*datasets* contain the processed data from the People Survey aligned to the
unique identifiers while *lookups* provide labels and other metadata relating
to the unique identifiers.

The harmonised datasets structure the data in a 'long format', that is each
row in a dataset relates to an individual value/score from the People Survey,
at a minimum each dataset includes the following columns:

* `data_type`: a value representing the type of data included in the dataset
  (see the [data typology](#sec-intro-typology) section for more details)
* `year`: the year of the survey's results the data relates to
* `uid_qm_num`: the unique numeric identifier relating to the question or
  measure
* `uid_qm_txt`: the unique text identifier relating to the question or measure
* `response_category`: a general identifier relating to the response category
  that the value represents (see the
  [response cateogries](#sec-intro-response-categories) section for more
  details).
* `value`: the actual survey result/score ranging from 0 to 100 (up to three
  decimal points). Scores typically relate to percentages (e.g. 23 would
  represent a true value of 23%, or 0.23 in decimal notation).

Datasets may also include identifiers relating to the organisation
(`uid_org_txt`) and/or demographic objects (`uid_demcat_num`, `uid_demcat_txt`,
`uid_demq_txt` and/or `uid_cat_txt`) that the data relates to.

```{r}
#| echo: false
set.seed(215202)
```

```{r}
#| label: tbl-bench-example
#| tbl-cap: Example structure of the benchmark results file
readr::read_csv(
    "../data/01-benchmarks/csps_benchmarks_2009-2024.csv",
    show_col_types = FALSE
) |>
    dplyr::sample_n(10)
```

[cspshub]: https://www.gov.uk/government/collections/civil-service-people-survey-hub