# Utilities and helper functions

A number of utilities and helper functions have been written to help the
processing of the People Survey data and are documented in this chapter.
The scripts for these utilities and helpers are contained in the
[`R/utils`]({{< var repo_links.r_utils >}}) folder of the `csps-data` repo.

## `csps_data_files`

There are a large number of files that provide the
[source data](03_source_data.qmd) for the People Survey. The
`raw-data/00_data_files.yaml` file provides a log (in YAML format) of all the
source data files stored in the `raw-data` folder of the `csps-data` repo.

The `R/utils/data_files.R` script loads this YAML file as a list
(called `csps_data_files`) to make it easier to access file paths from within
the R code.

```{r}
#| include: false
knitr::opts_knit$set(root.dir = "..")
```

```{r}
source("R/utils/data_files.R")
names(csps_data_files)
str(csps_data_files$csps2015)
```

The `R/data_files_ref.R` script also processes the YAML file to produce two
reference CSV files, it is not directly used in the processing.

## Data extraction helpers

The `R/utils/data_extract_helpers.R` function provides a series of helper
functions used in the scripts in folder `R/04-extract_data` that are designed
to open and process the raw data_files.

While the file formats of published data have changed over time (from CSV and
Excel to ODS format) and the internal structure of each format is different,
for each file type the structure has remained largely consistent throughout
time (for example the CSV of benchmark scores in 2018 is similar in structure
to the CSV of benchmark scores in 2013). This stability in per-format structure
has made it easy to construct a set of re-usable extraction functions:

* [`long_csv()`](#long_csv)
* [`extract_benchmark_ods()`]()
* `extract_organisation_ods()`
* `extract_demographic_data()`

All of these functions are designed to return a
[tibble](https://tibble.tidyverse.org/reference/tibble.html) that stores the
data from the raw data sets in 'long' format, that is each row of the tibble
relates to an individual value (i.e. a cell of the 2-dimensional input
spreadsheet/table).

```{r}
#| echo: false
#| label: fig-pivot-long
#| fig-cap: Example of converting People Survey data from 'wide' to 'long' format

inlinesvg::inline_svg_file(
    "companion/images/pivot_plot.svg",
    alt_title = "Example of a table being pivoted",
    alt_description = paste(
        "Image showing two tables. The left hand table shows an example of",
        "a People Survey benchmark scores dataset in 'wide' format with",
        "one question per row and one year per column. The right hand table",
        "shows the same data in 'long' format where each row is a separate",
        "value and the column header years have been transposed to a new",
        "column called year."
    ),
    dimensions = "strip"
)
```

### `long_csv()` {#long-csv}

The `long_csv()` function reads a CSV file and converts it into long format.
It is used to process CSV files containing either benchmark scores (and all
respondent scores which have the same layout) or organisation scores datasets.
Benchmark scores datasets have question labels in rows and years in columns,
while organisation scores datasets have organisation names (and, depending on
year, reference codes) in rows and question labels in columns.

```r
long_csv(
    path, type = NULL, values_convert = NULL, skip = 0, cols = NULL,
    na = c("", "NA", "n/a"), ...
)
```

path
: file path

type
: type of CSV being processed (either `benchmark_csv` or `organisation_csv`)

values_convert
: if necessary use `"scale_100"` to convert percentages from 0-1 decimal to
  0-100 score

skip
: rows to skip

cols
: columns to subset to

na
: values to consider missing

...
: values to pass on to `readr::read_csv()`

### `extract_benchmark_ods()`

The `extract_benchmark_ods()` function 

## `extract_response_cat()`

For the majority of attitudinal questions and measures in the People Survey the
published data includes only one value. The majority of attitudinal questions
are asked on a 'strongly agree' to 'strongly disagree' scale, and published
score is the combined percentage responding 'agree' or 'strongly agree'.
However, some questions use other scales and in some cases more than one
response category is published. There are also a small number of questions
with multiple responses.

The `extract_response_cat()` function (in the `R/utils/extract_response_cat.R`
file) takes a vector of attitudinal question/measure text and associated UIDs
and returns a vector of response categories based on that input.